Geometric models reveal behavioral and neural signatures of transforming experiences into memories

### 2020-11-24 paper/main.tex

\begin{abstract}
The mental contexts in which we interpret experiences are often person-specific, even when the experiences themselves are shared. We developed a geometric framework for mathematically characterizing the subjective conceptual content of dynamic naturalistic experiences. We model experiences and memories as ``trajectories'' through word embedding spaces whose coordinates reflect the universe of thoughts under consideration. Memory encoding can then be modeled as geometrically preserving or distorting the ``shape'' of the original experience. We applied our approach to data collected as participants watched and verbally recounted a television episode while undergoing functional neuroimaging. Participants’ recountings all preserved coarse spatial properties (essential narrative elements), but not fine spatial scale (low-level) details, of the episode’s trajectory. We also identified networks of brain structures sensitive to these trajectory shapes. Our work provides insights into how we preserve and distort our ongoing experiences when we encode them into episodic memories.
\end{abstract}

>>>>

\begin{abstract}
How do we preserve and distort our ongoing experiences when encoding them into episodic memories?
The mental contexts in which we interpret experiences are often person-specific, even when the experiences themselves are shared. We developed a geometric framework for mathematically characterizing the subjective conceptual content of dynamic naturalistic experiences. We model experiences and memories as ``trajectories'' through word embedding spaces whose coordinates reflect the universe of thoughts under consideration. Memory encoding can then be modeled as geometrically preserving or distorting the ``shape'' of the original experience. We applied our approach to data collected as participants watched and verbally recounted a television episode while undergoing functional neuroimaging. Participants’ recountings all preserved coarse spatial properties (essential narrative elements), but not fine spatial scale (low-level) details, of the episode’s trajectory. We also identified networks of brain structures sensitive to these trajectory shapes.
\end{abstract}
